A hazards approach to the biometric analysis of infant mortality

### 2023-03-21 paper2.Rmd
work on tabular data, not completed

Note the two outliers, February and October. Born in February means a tough first month of 
life, but surviving that, the future prospects are great. The opposite holds for those born in October.

@@

```{r plotperiod2, fig.cap = "By calendar season, periods.", fig.height = 8, eval = FALSE}
oldpar <- par(mfrow = c(2, 2))
for (seas in levels(dat$season)){
    drawBB2(dat[dat$season == seas, ], main = seas, printLegend = FALSE)
}
par(oldpar)
```
The periods are of course more homogeneous regarding weather and external threats,
and it shows off in that in all four periods the suggested models fit well, here shown by the
"constant hazards model" for exogenous mortality.

@@

The assumptions about constant hazard and uniform distribution of exogenous mortality 
are only applicable on population level, it can never hold for an individual. It can 
be thought of as an average over birth cohorts, where

>>>>

Note the two outliers, February and October. Born in February means a tough first month of 
life, but surviving that, the future prospects are great. The opposite holds for those born in October, 
no problems during the neonatal phase, but then comes the tough winter months.

@@

```{r plotperiod2, fig.cap = "By calendar season, periods.", fig.height = 8, eval = FALSE}
oldpar <- par(mfrow = c(2, 2))
for (seas in levels(dat$season)){
    drawBB2(dat[dat$season == seas, ], main = seas, printLegend = FALSE)
}
par(oldpar)
```
The periods are of course more homogeneous regarding weather and external threats,
and it shows off in that in all months the suggested models fit well, here shown by the
"constant hazards model" for exogenous mortality.

@@

The assumptions about constant hazard and uniform distribution of exogenous mortality 
are only applicable on population level, it can never hold for an individual. It can 
be thought of as an average over birth cohorts, where it is important to have birth dates 
spread evenly within calendar years.

## Tabular data

Often the Bourheois-Pichat method is applied to tabular data. The most extreme case is
that the only available data are number of births $B$, number of neonatal deaths $D_1$, and total 
number of deaths $D$. The Bourgeois-Pichat line fitting is very simple to do in this case, but 
how about the hazards approach?

It turns out that it is easy to find given the data at hand followíng the procedure given in 
Section 4.3: The probability of dying in the post-neonatal period, given survival of the neonatal
period is estimated by $hat{p} = (D-D_1) / (B - D_1)$, and the length of the post-neonatal time span 
is $g(365) - g(28) \approx 297$ "g-units".
